---
title: "Incarceration Across King County, Washington"
output: html_document
---

```{r setup, include=FALSE, echo=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("analysis.R")
source("trends_graph.R")
source("var_compar.R")
source("map.R")
```
---

### Introduction:

The primary incarceration data and trends used for this project was gathered by the Vera Institute of Justice from two primary sources: the Bureau of Justice Statistics and the Annual Survey of Jails. This data has since been manipulated and reduced to showcase underlying trends within
King County, Washington. 

King County was selected for this project for its  diversity. While it varies depending on the location within the county, the general diversity and associated incarceration trends can serve as an example for what might found in other U.S. metropolitan areas. 

---

### Summary Information:

To start, I analyzed ratios of incarcerated populations of a given racial demographic to their overall population in King County in the year 2018. I wanted to see how inconsistent the proportion of incarcerated people would be across different racial groups in a fairly diverse county, especially one that preaches its progressiveness. In doing so, I found that among Asian Americans and Pacific Islanders, the ratio was `r analysis$recent_aapi_ratio`. For the Latinx community, it was `r analysis$recent_latinx_ratio`. Indigenous population, `r analysis$recent_native_ratio`. White population, `r analysis$recent_white_ratio`. For the Black community, the ratio was `r analysis$recent_black_ratio`. 

The ratio of Black people in incarceration to the total Black population in King County was nearly five times that of the White population in King County, and 17 times that of the smallest ratio. In a fair society, we would expect these ratios to be fairly similar across the board. However, the United States continues to grapple with structural racism and inequities in its integrity. 

These ratios have fluctuated slightly over the last 30 years, but they remain lie in the same area. The ratio of incarcerated Black people in King County remains the highest, followed by that of incarcerated Indigenous people, then Latinx, White, and AAPI. 

In attempts to adjust for confounding variables, I pulled  ratios of the incarcerated racial group in King County to the total incarcerated population in each given year. The following ratios were `r analysis$recent_ratio_aapi_of_jail_pop` for AAPI's in King County, `r analysis$recent_ratio_black_of_jail_pop` for the Black community, `r analysis$recent_ratio_latinx_of_jail_pop` for the Latinx community, `r analysis$recent_ratio_native_of_jail_pop` for Indigenous communities, and `r analysis$recent_ratio_white_of_jail_pop` for the White population. We see that the some ratios level out and become closer to that of other communities, but still, there are drastic disparities across other racial groups that we cannot assume are naturally occurring. Marginalization, racial profiling, and unequal distribution of resources are just a few things that contribute to these varying ratios. With new information being shared about these issues, it is imperative that our society, from its leadership to its citizens, truly recognizes them as issues and works to address them. 

--- 

### Showcasing Trends:
```{r echo = F}
trends_graph
```
This graph shows the changes in ratios of an incarcerated racial group to its entire population in King County. It's fair to assume none of the trend lines will ever be straight across, but it's interesting to see how they change over the years. Some linger in the same range while others show peaks that coincide with that of other racial groups. By showcasing the trend lines on one graph, we can see how they differ from each other and how they change over time. It raises questions of why some peak and others don't, and what events those peaks align with. These are questions that have yet to be answered by data, and perhaps with further research, we can come to an answer. 

For example, the ratios for AAPI and White communities are generally stagnant and are consistent throughout the years. However, mng Black, Indigenous, and Latinx communities, those ratios are significantly higher and experience peaks together around the mid-1990's, early 2000's, and early 2010's. These peaks may coincide with times of economic distress or societal disorder. Again, further research must be conducted, but we hope to find answers to illuminate these areas of question. 

--- 

### Showcasing Relationships:
```{r echo = F}
var_compar_graph
```

The population to incarcerated population graphs for AAPI, Black, Indigenous, and White communities in King County all show a gradual linear increase. This is fair to expect, because as the population grows, the incarcerated population might increase with it. However, for the Latinx community, just as the population reaches 50,000 people, the incarcerated population jumps four-fold, and then seven and nine-fold before the population even grows to 75,000.  These are such large jumps, and they are only characteristic to the Latinx community in King County. It makes me wonder what is going on, because I can't imagine this occurring organically. 

---
### Map:
```{r echo = F}
incar_map
```

Featured is a chart of Washington state with all of its counties. The color scale refers to the total incarcerated population in each county. The larger the population, the darker the color. As we can see, King County has the highest count of incarcerated people. While this is alarming, it should be recognized that King County is the largest county in the state. Moving outside of King County, the county population drops, and so does the the number of people in jails. 



